---
title: "Example of fitting incidence-mortality model"
author: "Jeff Eaton"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is a working document to illustrate the basic tools of the analysis.

## Load data 

Prepped cohort datasets are stored separately in the `alphaepisdata` package. Load this package to access the data sets.

```{r load data}
devtools::load_all("~/Documents/Code/R/alphaepisdata", export_all=FALSE)
devtools::load_all()

## Load individual, residences, and HIV tests datasets
data(alpha_ind, alpha_res, alpha_hiv)

hiv <- subset(alpha_hiv, type %in% c("survey", "self report") & !is.na(result) & !is.na(testdate))
res <- alpha_res
ind <- alpha_ind

```

The `prepare.stan.data()` function sets up stan input data for a model fit. Currently, 
the function uses datasets from the global environment, among other greivances which 
should be improved.

```{r}
rakaif_stand <- prepare.stan.data(c('Rakai', '1975prior451'), 'Female',
                                  min.time=1970, max.time=2015, min.age=15, max.age=100,
                                  natmxstart.time=1994, artstart.time=2004.5, cohortstart.time=1970, cohortend.time=2015,
                                  pen.ord.incrate=1, pen.ord.natmx.time=1, pen.ord.natmx.age=1, pen.ord.art=1,
                                  dt=0.2, k.dt=5, nk.art=5, hivonly=FALSE, hivelig=FALSE)
```

Call `rstan::sampling()` to fit the model. Usually, I run approximately 4 chains for 500 iterations. This is best done on a cluster.

```{r, fit model}
fit <- rstan::sampling(alphaepi::stanmodels$le_interact, data=rakaif_stand,
                       iter=25, chains=1, chain_id=1, refresh=1)
```

The approach to analysis of model outputs is to create an output object and then
call functions to analyse posterior outputs.

```{r, model outputs}
mod <- list(fit = fit, stand = rakaif_stand)
mx_out <- add.mx(mod)

saveRDS(mod, "fit-example/mod.rds")
```


## Likelihood calculation

0. A set of parameter inputs

```{r}
param <- create.param.list(mod$fit)
length(param) # number of resamples
names(param[[13]]) # each has a single sample from joint posterior

p <- param[[13]]
stand <- mod$stand
``` 

1. Construct arrays for rates

* $\lambda(t, a)$: incidence rate by time and age
* $\mu(t, a)$: non-HIV mortality by time and age

Note: `create.modpred()` is a magic function which illustrates many of these steps in an undocumented way.

```{r}
incrateMID_time_age <- exp(stand$Xmid_incrate_time %*% p$coef_incrate_time_age %*% t(stand$Xmid_incrate_age))
natmx_time_age <- exp(stand$X_natmx_time %*% p$coef_natmx_time_age %*% t(stand$X_natmx_age))
```

2. Calculate cumulative probability of avoiding HIV and survival by cohort

```{r}
cumavoid_time_age <- exp(-stand$dt*diagCumSum(incrateMID_time_age))
cumavoidMID_time_age <- cumavoid_time_age[1:(stand$STEPS_time-1), 1:(stand$STEPS_age-1)]*exp(-stand$dt/2*incrateMID_time_age)
natsurv_time_age <- exp(-stand$dt*diagCumSum(exp(stand$Xmid_natmx_time %*% p$coef_natmx_time_age %*% t(stand$Xmid_natmx_age))))
```


```{r}
m <- create.modpred(p, stand)

calc_phivn(37, 52, death = 0, hivpos = 0, m$cumavoid_time_age, m$natsurv_time_age, m$natmx_time_age)

## If hivpos = 1, then returns 0 since zero probability that I remained HIV negative
calc_phivn(37, 52, death = 0, hivpos = 1, m$cumavoid_time_age, m$natsurv_time_age, m$natmx_time_age)
```


Function `calc_phivsurv()` calculates the probaility of survival for each time step of exposure. The
below calculates someone who tested negative at (time, age) = (10, 15), tested positive at 
(time, age) = (22, 27), and survived to (time, age) = c(32, 37).

```{r}

calc_phivsurv(tIDX = 10,       # time step of exposure start (last time known negative)
              aIDX = 15,       # age index of exposure start
              exposeDUR = 12,  # time steps of exposure
              exit_tIDX = 32,  # censor time step
              m$hivsurv_dur_a0,    # array: probability of surviving by duration and age at infection 
              m$hivmxMID_dur_a0,   # array: instantaneous mortality probability at midpoint of time step
              m$artrr_MID,         # vector: reduction in HIV mortality by time step
              stand$artstart_tIDX, # integer: time step when ART starts
              stand$dt)            # real: time step duration
```

This has returned a vector of length 12 corresponding to the probability of surving to `t = 32` given 
infection occurred in time step `s0 = 10:21`, respectively. The vector is increasing because the probability
I survive to `t = 32` is greater if infection occurred later (shorter duration of infection between `s0` and `t=32`.

The function `calc_phivp()` calculates the probability of surviving to exit time given the relevant exposure interval.
This integrates the cumulative incidence and survival distributions over the exposure interval.
```{r}
calc_phivp(exit_tIDX = 32,
           exit_aIDX = 37,
           expose_tIDX = 10,
           expose_aIDX = 15,
           expose_DUR = 12, 
           death = 1,
           m$cumavoidMID_time_age,
           m$incrateMID_time_age,
           m$hivsurv_dur_a0, 
           m$hivmx_dur_a0,
           m$hivmxMID_dur_a0,
           m$artrr,
           m$artrr_MID,
           stand$artstart_tIDX, 
           m$natsurv_time_age,
           m$natmx_time_age,
           stand$dt)
```
